Spectral Variability of Two Rapidly Rotating Brown Dwarfs: \\2MASS~J08354256$-$0819237 and 2MASS~J18212815+1414010

### 2017-02-14 ms.tex
I wiped out the appendix b/c it's now split into a supplement document.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j1821_t_diff.pdf}
\caption{The same data as in Figure \ref{fig:ampspec1821mie}. We fit the data with a two region model that assumes the effective temperature of the brown dwarf and a cooler temperature.}\label{fig:ampspec1821tdiff}
\end{centering}
\end{figure}
% Command used:
% plot_rad_vs_wavl,/amp,/shadetell,custxrange=[0.8,2.5],custyrange=[-0.1,1.9],/showYang,/twoTreg

@@

\appendix
\clearpage
\pagebreak

\section{Color-Magnitude Diagram}

Figure \ref{fig:CMD} shows the color magnitude diagram from the sample of brown dwarfs in \citet{dupuy2012ltparallax}, which includes 2MASS J0835 but not 2MASS J1821.
We highlight these two brown dwarfs to be contrasted with the rest of the population.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{color_mag.pdf}
\caption{Color-Magnitude diagram from \citet{dupuy2012ltparallax}. We added 2MASS J1821 using Simbad magnitudes and parallaxes.}\label{fig:CMD}
\end{centering}
\end{figure}

\pagebreak
\section{Example Posterior Distributions}
[DO WE NEED MORE TEXT IN THIS SECTION EXPLAINING EACH FIGURE?]
\subsection{Time Series}
In Figure \ref{fig:postCosfit}, we show the posterior distributions of a sinusoidal model (Equation \ref{eq:cosfit}) fit to the data using \texttt{emcee} \citep{foreman-mackey2013emcee} for an example wavelength bin of 1.08~$\mu$m.
This wavelength bin is one of the high amplitude fits that is relatively unaffected by telluric contamination.
In this example, we input photon noise only.
We impose priors on the amplitude $A_1 > 0$, $|t_t| < 6 hr$ (to avoid multiple periodic solutions) and that the period $0 < \tau < 6 hr$ so that the periodicity is shorter than the observation window length.
The model parameters are close to Gaussian and the only visible correlation is between the time offset, $t_1$, and the rotation period $\tau$.
The maximum likelihood solution is shown in Figure \ref{fig:model2Cosfit} and has a chi-squared per degree of freedom ($\bar{\chi^2}$) of 12.1

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{corner_example.pdf}
\caption{Example posterior distributions for an MCMC fit to the time series using Equation \ref{eq:cosfit} for the wavelength bin 1.08~$\mu$m.
Amplitude, $A_1$ is in percent, the time offset $t_1$ is in hours, the period of oscillation, $\tau$, is in units of hours, B is in units of fraction of flux per hour and C is the offset constant in fraction (not percent).
We only input photon counting and read noise statistics for the errors.
This is likely and underestimate for the uncertainty in the time series.
}\label{fig:postCosfit}
\end{centering}
\end{figure}

Since Equation \ref{eq:cosfit} is a poor fit to the data, as visible in Figure \ref{fig:model2Cosfit}, we also explore models with additional sinusoidal frequencies.
In the multiple term model, we add cosine functions at $n$ times the frequency:
\begin{equation}\label{eq:cosfitMultiTerm}
F(t) = B t + C + A_1 \cos(2 \pi (t - t_1)/\tau) + A_2 \cos(4 \pi (t - t_2)/\tau) + A_n \cos(2 n \pi (t - t_n)/\tau) +  ...
\end{equation}
Now there are $n$ amplitude parameters $A_n$ and $n$ time offset parameters $t_n$.
Figure \ref{fig:model2Cosfit} shows the best fit compared to the data for 2 term and 3 term sinusoidal fits.
We assume the same priors as the 1 term sinusoidal fit with  $A_n > 0$.
For the time offsets, we restrict the range to be $\pm$ 4 hr / $n$.
The multiple term model appear visually to fit the data well.
However, the $\bar{\chi^2}$ are still 7.9 and 7.4, so it is not the expected 1.0.
However, the error bars are underestimated when only including photon and read noise components.
With exoplanet light curves, the standard deviation of the out-of-transit baseline flux can be used as an estimate for errors. With the brown dwarf rotation light curves, there is no flat baseline over which to measure error levels.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{best_fit_multi_terms.pdf}
\caption{A zoom-in on on the time series of 2MASS J1821 in Figure \ref{fig:tserPhase1821} for the wavelength bin 1.08~$\mu$m. Error bars are shown for photon and read noise. We plot a best-fit models using a single cosine term (green, $\bar{\chi^2}$ = 12.1). Additional terms improve the fits to the data and the $\bar{\chi^2}$.}\label{fig:model2Cosfit}
\end{centering}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.7\textwidth]{corner_fit_2term.pdf}
\caption{Example posterior distributions for an MCMC fit to the time series using Equation \ref{eq:cosfitMultiTerm} for 2 terms.
Again, there is little correlation between the posteriors for the different parameters. {\it Right} Same figure, but with 3 cosine terms.}\label{fig:post2Cosfit}
\end{centering}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.48\textwidth]{residual_2term_1080nm.pdf}
\includegraphics[width=0.48\textwidth]{residual_3term_1080nm.pdf}
\caption{Residuals after subtracting the maximum likelihood two-term model from Equation \ref{eq:cosfitMultiTerm} from the data.  We show the residuals when using a 2 term model (left) and a 3 term model (right). The standard deviation of the residuals is 0.6\% whereas the error from photon counting statistics is 0.2\%.}\label{fig:resid2Cosfit}
\end{centering}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{histo_resids.pdf}
\caption{Histogram of residuals after subtracting the maximum likelihood model from Equation \ref{eq:cosfitMultiTerm} from the data with 2 terms.}\label{fig:histoResid2Cosfit}
\end{centering}
\end{figure}

\clearpage
\pagebreak
\subsection{Amplitude Spectrum}

We also use \texttt{emcee} to fit the amplitude spectrum of 2MASS J1821 using the two temperature surface model in Section \label{sec:spotModel} with Equation \ref{eq:twoTempAmp}.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{best_fit_spec_2temp.pdf}
\caption{We repeat the amplitude spectrum of 2MASS J1821 from Figure \ref{fig:ampspec1821tdiff}, which excludes the wavelengths contaminated by telluric absorption.
We show a maximum likelihood model spectrum, which has a $\chi^2$ per degree of freedom of 1.16.}\label{fig:ML2Tempfit}
\end{centering}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{corner_fit_2temp.pdf}
\caption{Example posterior distributions for an MCMC fit to the spectra using Equation \ref{eq:twoTempAmp}.}\label{fig:post2Tempfit}
\end{centering}
\end{figure}

\clearpage
\pagebreak
\subsection{Manual Shifting}

The trends near 1.4 $\mu$m and 1.8 $\mu$m in \shb\ could probably be reduced by better wavelength shifting.
I wrote a GUI that let me control the shifts for each spectrum and plot the ratio spectrum.
I minimized the residuals and ran a test case for \sha.
As seen in Figure \ref{fig:manualVsAutoShift0835}, the manual shifting dramatically reduces the systematic linear trends in the dynamic spectra.
In contrast to Figure \ref{fig:specphot}, not linear de-trending has been applied to columns so the systematic linear trends are visible in both cases.
The automatic shifting finds the peak in the cross correlation.
It does this by cross-correlating a portion of the spectrum (excluding the endpoints) and zeroed on the wings. 
The automatic method then fits the cross-correlation pattern with a parabola to find the peak.
The manual method is where we use keystrokes to shift one spectrum by 0.333 px back in forth until the ratio of another spectrum has the smallest residuals.
Two iterations were used in the manuals shifting: (1) minimizing the residuals of the target divided by reference and (2) minimizing the residuals of the target divided by the target middle spectrum.

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835- Automatic Shifting]{
	\includegraphics[width=0.5\textwidth]{specphot_0835_auto_noremovel.pdf}
	\label{fig:specphot0835autoS}
	}
\subfloat[2MASS J0835 - Manual shifting ]{
	\includegraphics[width=0.5\textwidth]{specphot_0835_manual_noremovel.pdf}
	\label{fig:specphot0835manualS}
	}
	\caption{Dynamic spectra of {\sha} using automatic cross correlation wavelength shifting (left) and manual wavelength shifting(right).
	The manual method shows dramatically better systematics in these dynamic spectra which have no fits along columns to remove trends.}
	\label{fig:manualVsAutoShift0835}
	\vspace{0.1in}
\end{figure*}

\begin{figure*}[!t]
\centering
\subfloat[2MASS J1821- Automatic Shifting]{
	\includegraphics[width=0.5\textwidth]{specphot_1821_auto_noremovel.pdf}
	\label{fig:specphot1821autoS}
	}
\subfloat[2MASS J1821 - Manual shifting ]{
	\includegraphics[width=0.5\textwidth]{specphot_1821_manual_noremovel.pdf}
	\label{fig:specphot1821manualS}
	}
	\caption{Dynamic spectra of {\shb} using automatic cross correlation wavelength shifting (left) and manual wavelength shifting(right).
	The manual method shows better systematics but higher frequency noise.}
	\label{fig:manualVsAutoShift1821}
	\vspace{0.1in}
\end{figure*}

\clearpage
\pagebreak
\subsection{MCMC Time Series Fits - first try}

Fitting all light curves to \sha\ with MCMC methods and without the A$_1 >$ 0 prior results in more non-detections, but an unusual pattern to the spectra.
As with other MCMC fits, error bars are from photon and read noise, so they're likely under-estimated.

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835 - Manual Shifting]{
	\includegraphics[width=0.5\textwidth]{mcmc_spectrum_2mass_0835.pdf}
	\label{fig:ampSpec0835ManualS}
	}
\subfloat[2MASS J1821 - Manual shifting ]{
	\includegraphics[width=0.5\textwidth]{mcmc_spectrum_2mass_1821.pdf}
	\label{fig:ampSpec1821ManualS}
	}
	\caption{Amplitude spectrum of {\sha} (left)  and {\shb} (right) using manual wavelength shifting, 1 cosine term fit and MCMC fitting with error bars from photon and read noise.}
	\label{fig:ampSpecManualS}
	\vspace{0.1in}
\end{figure*}

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835 - 4$\times \sigma$]{
	\includegraphics[width=0.5\textwidth]{mcmc_2mass_0835_spectrum_0400perc_err.pdf}
	\label{fig:ampSpec0835fourTimesErr}
	}
\subfloat[2MASS J1821 - 4$\times \sigma$ ]{
	\includegraphics[width=0.5\textwidth]{mcmc_2mass_1821_spectrum_0400perc_err.pdf}
	\label{fig:ampSpec1821fourTimesErr}
	}
	\caption{As an experiment, I remade the amplitude spectra of Figure \ref{fig:ampSpecManualS} but scaled all error bars up by a factor of 4 for all points to account for some systematic errors. IE, Amplitude spectrum of {\sha} (left)  and {\shb} (right) using manual wavelength shifting, 1 cosine term fit and MCMC fitting with error bars from 4$\times$ photon and read noise.}
	\label{fig:ampSpecManualSfourTimesErr}
	\vspace{0.1in}
\end{figure*}

\clearpage
\pagebreak
\subsection{MCMC Time Series Fits - second try}

As with my \texttt{IDL} code, I found that I had to impose a tight prior on the rotation period (I actually fixed it at the literature value of 4.1 hr) in order to avoid fitting the systematics at the telluric wavelengths.
When I do this, the MCMC code produces a very smooth amplitude spectrum for 2MASS 1821, as visible in Figure \ref{fig:ampSpecManualSTauRestrict}.
I experimented with the restriction that $3.8 < \tau < 4.0$ based on the multi-term Fourier series fits.
I used two different error scalings - 2.5 $\sigma$ and 4 $\sigma$.
2.5 $\sigma$ give the expected $\bar{\chi^2} \approx 1$, whereas 4 $\sigma$ gives around 0.4.

\begin{figure*}[!t]
\centering
\subfloat[2MASS J1821 - 2.5$\times \sigma$]{
	\includegraphics[width=0.5\textwidth]{2mass_1821_spectrum_0250perc_err_taurestrict.pdf}
	\label{fig:ampSpec1821times2p5ErrTauRest}
	}
\subfloat[2MASS J1821 - 4$\times \sigma$ ]{
	\includegraphics[width=0.5\textwidth]{2mass_1821_spectrum_0400perc_err_taurestrict.pdf}
	\label{fig:ampSpec1821fourTimesErrTauRest}
	}
	\caption{Restricting the $\tau$ parameter to, $3.8 < \tau < 4.0$ significantly reduces the long term quadratic terms' contribution to the sinusoidal fit. In this case, there were MCMC fits for \shb, one with error bars scaled by 2.5$\sigma$ and the other by 4.0$\sigma$.}
	\label{fig:ampSpecManualSTauRestrict}
	\vspace{0.1in}
\end{figure*}

As with \shb\ I imposed a tight prior on $\tau$ such that $3.0 < \tau < 3.2$ for \sha. This now gives an average amplitude of zero -- see Figure \ref{fig:ampSpec0835TauRestrict}.

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835 - 2.5$\times \sigma$]{
	\includegraphics[width=0.5\textwidth]{2mass_0835_spectrum_0250perc_err_taurestrict.pdf}
	\label{fig:ampSpec0835times2p5ErrTauRest}
	}
\subfloat[2MASS J0835 - 4$\times \sigma$ ]{
	\includegraphics[width=0.5\textwidth]{2mass_0835_spectrum_0400perc_err_taurestrict.pdf}
	\label{fig:ampSpec0835fourTimesErrTauRest}
	}
	\caption{Restricting the $\tau$ parameter to $3.0 < \tau < 3.2$ results in a spectrum that is on average near 0\% amplitude for \sha. We show the amplitude spectra for error bars scaled by 2.5$\sigma$ and 4.0$\sigma$.}
	\label{fig:ampSpec0835TauRestrict}
	\vspace{0.1in}
\end{figure*}

One thing to check is the time offset parameter as a function of wavelength.
AB noticed a trend in this $t_1$ parameter, but didn't have error bars fully incorporated.
I'm finding the trend is marginal when using 4$\sigma$ noise scaling.

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835 - 4$\times \sigma$]{
	\includegraphics[width=0.5\textwidth]{2mass_0835_t_1_spectrum_0400perc_err_taurestrict.pdf}
	\label{fig:t1Spec0835TauRest}
	}
\subfloat[2MASS J1821 - 4$\times \sigma$ ]{
	\includegraphics[width=0.5\textwidth]{2mass_1821_t_1_spectrum_0400perc_err_taurestrict.pdf}
	\label{fig:t1Spec1821TauRest}
	}
	\caption{Searching for trends in the $t_1$ parameter as a function of wavelength. The possible upward trend is marginal for \shb\ when using the 4$\sigma$ scaling of error bars. In these cases, we have restricted the $\tau$ parameter with priors to $3.0 < \tau < 3.2$ and $3.8 < \tau < 4.0$ for \sha\ and \shb\ respectively.}
	\label{fig:t1SpecTauRestrict}
	\vspace{0.1in}
\end{figure*}

\begin{figure*}[!t]
\centering
\subfloat[2MASS J0835 - 4$\times \sigma$]{
	\includegraphics[width=0.5\textwidth]{2mass_0835_t1_vs_A1_0400perc_err_taurestrict.pdf}
	\label{fig:t1vsA1Spec0835}
	}
\subfloat[2MASS J1821 - 4$\times \sigma$ ]{
	\includegraphics[width=0.5\textwidth]{2mass_1821_t1_vs_A1_0400perc_err_taurestrict.pdf}
	\label{fig:t1vsA1Spec1821}
	}
	\caption{Searching for trends in the $t_1$ parameter as a function of Amplitude ($A_1$). No clear trend arises in these correlation plots. As with Figure \ref{fig:t1SpecTauRestrict}, we have used 4$\sigma$ scaling of error bars and restricted the $\tau$ parameter with priors to $3.0 < \tau < 3.2$ and $3.8 < \tau < 4.0$ for \sha\ and \shb\ respectively.}
	\label{fig:t1vsA1SpecTest}
	\vspace{0.1in}
\end{figure*}

\clearpage
\pagebreak
\subsection{Mie scattering}

My old Mie scattering code was in IDL and would be more work to incorporate into MCMC fitting, so I found python package \texttt{miescatter} and did a comparison with my IDL code.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{python_vs_IDL_example.pdf}
\caption{Example Mie scattering extinction cross sections $Q_{ext}$ for a spherical particle with a radius of 0.5~$\mu$m and a complex index of refraction of n=1.825 - 0.001 $i$. The IDL and python code's are in excellent agreement.}\label{fig:IDLvsPython}
\end{centering}
\end{figure}

I did an initial MCMC fit where I plugged in these same index of refraction (n=1.825 - 0.001 $i$ for all wavelengths) and fit the spectrum that I found in IDL.
The best fit matches the data well with a $\bar{\chi^2}$ of 1.3, visible in Figure \ref{fig:mcmcMieBest1particle}.
However, there are many solutions at different resonances in the Mie extinction curve visible in Figure \ref{fig:mcmcMieCorner1particle}.
A better way to handle this is a log-normal distribution of particles as in \citet{2016ApJ...826..156S}.

\begin{figure*}[!t]
\centering
\subfloat[Single Particle Best fit for \shb]{
	\includegraphics[width=0.5\textwidth]{best_fit_single_particle_mie.pdf}
	\label{fig:mcmcMieBest1particle}
	}
\subfloat[Single Particle Posterios for \shb]{
	\includegraphics[width=0.5\textwidth]{corner_single_particle_mie.pdf}
	\label{fig:mcmcMieCorner1particle}
	}
	\caption{An initial fit to the amplitude spectrum when using a single particle.
	The best fit nicely matches the data but has multiple solutions at the Mie scattering resonances.}
	\label{fig:mcmcMieCorner1particle}
	\vspace{0.1in}
\end{figure*}

I manually created a log-normal distribution of particles in Python and added up their cross sections weighted by number of particles.
This process is slow, so I calculated it once and found  a 15th order polynomial fit to $Q_{ext}$ and evaluated this in the model.
The resulting MCMC distribution is shown in Figure \ref{fig:mcmcMieCornerLogNorm}.

\begin{figure*}[!t]
\centering
\subfloat[Log-Normal Best fit for \shb]{
	\includegraphics[width=0.5\textwidth]{best_fit_lognormal_mie.pdf}
	\label{fig:mcmcMieBestLogNorm}
	}
\subfloat[Log-Normal Posteriors for \shb]{
	\includegraphics[width=0.5\textwidth]{corner_lognormal_mie.pdf}
	\label{fig:mcmcMieCornerLogNorm}
	}
	\caption{An initial fit to the amplitude spectrum when using a log-normal size distribution with a $\sigma=0.5$.
	Again, there are a few ``walkers'' that stay at large radii.}
	\label{fig:mcmcMieCornerLogNorm}
	\vspace{0.1in}
\end{figure*}

On closer inspection, I see that there is a local minimum for large particle radii where the $\chi^2$ begins to increase from 3~$\mu$m to 2$\mu$m.
This is illustrated in Figure \ref{fig:localMinLargeRadii}.
\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{local_min_illustration.pdf}
\caption{A set of example Mie scattering models for example parameters for the Amplitude offset and median particle radius written as [A,r]. The model at 2~$\mu$m is slightly worse than at 3~$\mu$m, so it traps a few walkers in the MCMC run.}\label{fig:localMinLargeRadii}
\end{centering}
\end{figure}

When I discard walkers with very low likelihood, as done in CHIMERA, the local minima are not included since their so low probability.
The resulting MCMC run is shown in Figure \ref{fig:localMinExcludeCorner}.

\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{corner_lognormal_mie_no_lmin.pdf}
\caption{MCMC fit when discarding points with a $\chi^2$ more than 5 times the minimum to exclude walkers that were trapped at local minima.}\label{fig:localMinExcludeCorner}
\end{centering}
\end{figure}

>>>>

\vspace{0.2in}
\begin{figure}
\begin{centering}
\includegraphics[width=0.5\textwidth]{amp_vs_wavl_j1821_t_diff.pdf}
\caption{The same data as in Figure \ref{fig:ampspec1821mie}. We fit the data with a two region model that assumes the effective temperature of the brown dwarf and a cooler temperature.}\label{fig:ampspec1821tdiff}
\end{centering}
\end{figure}
% Command used:
% plot_rad_vs_wavl,/amp,/shadetell,custxrange=[0.8,2.5],custyrange=[-0.1,1.9],/showYang,/twoTreg

@@
